Nonlinear methods to quantify Movement Variability in Human-Humanoid Interaction Activities

### 2020-10-28 docs/submission/arxiv/v02/main-arxiv.tex
axiv submission v02

%*******************************************************************************
\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\graphicspath{{../}} %goes to path: docs/

@@

\includegraphics[width=1.0\textwidth]{figures/methods/rss/pdf/rss}

@@

As an example of the use of $E_1(m)$ and $E_2(m)$ values, we consider two time 
series: the solution for the $x$ variable of the Lorenz system 
(Fig~\ref{fig:e1e2}E), and a Gaussian noise time series with zero mean 
and a variance of one (Fig~\ref{fig:e1e2}F).
We then compute $E_1(m)$ and $E_2(m)$ values for each time series.
The $E_1(m)$ values for the chaotic time series appear to be constant
after the dimension is equal to six.
The determination of six is given that any value of $m$ can be used as they 
are within the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}A).
$E_2(m)$ values, for chaotic time series, are different to one 
(Fig~\ref{fig:e1e2}C), for which, it can be concluded that for the 
chaotic time series the minimum embedding dimension the time series 
comes from a deterministic signal. With regard to the noise time series,  
$E_1(m)$ values appeared to be constant when $m$ is close to thirteen, 
which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}B).
$E_1(m)$ values then indicate the minimum embedding dimension of the 
noisy time series is thirteen, however all of the $E_2(m)$ values are 
approximately equal to one (Figure~\ref{fig:e1e2}D) for which it can be 
concluded that noise time series is a stochastic signal.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/methods/cao/pdf/cao}
    \caption{
	{\bf Minimum dimension embedding values with Cao's method.} 
	(A, B) $E_1 (m)$ values and (C, D) $E_2(m)$ values 
	with variations of $\tau$ values from one to twenty
	for (E) chaotic and (F) random time series.
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:e1e2}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

@@

For example, we compute the AMI for two time series:
A) the $x$ solution of the Lorenz system, and
B) a noise time series using a normal distribution with mean zero and 
standard deviation equal to one. From Fig~\ref{fig:amis}, it can then be 
concluded that the amount of knowledge for any noise time series is zero 
for which the first minimum embedding parameter is $\tau_0=1$. 
On the contrary, the first minimum of the AMI for the chaotic time series 
is $\tau_0=17$ which is the value that maximize the independence 
between $x(n)$ and $x(n+\tau)$ in the reconstructed state space 
\cite{bradley2015}.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/methods/ami/pdf/ami}
    \caption{
	{\bf Minimum delay embedding values with AMI's method.} 
    	(A, B) AMI values where its first minimum value in the curve
	is the minimum time delay embedding ($\tau_0$), 
	for (C) a chaotic and (D) noise time series.
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:amis}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------
Similarly as Cao's algorithm negatives, AMI's algorithm is not an
exception for negatives, which are worthwhile to mention for further 
investigations.
For instance, (i) is not clear why the choose of the first minimum of the AMI 
is the minimum delay embedding parameter \cite{kantz2003} and 
(ii) the probability distribution of the AMI function is computed
with the use of histograms which depends on a heuristic choice of number of bins
for which AMI depends on partitioning \cite{garcia2005e71}.

@@

For the creation of a recurrence plot based on time series 
$\{ \boldsymbol{x}_n \}$, it is first computed the state space 
reconstruction with uniform time-delay embedding 
$X(i)=\{ \boldsymbol{ \tilde{x} }_n, \dots,  
\boldsymbol{ \tilde{x} }_{n -(m-1)\tau} \}$
where $i=1,\dots,N$, $N$ is the number of considered states of $X(i)$ 
and $X(i) \in \mathbb{R}^m$ \cite{eckmann1987}.
The recurrence plot is therefore a two-dimensional $N \times N$ square 
matrix, $\mathbf{R}$, where a black dot is placed at $(i,j)$ 
whenever $X(i)$ is sufficiently close to $X(j)$: 
%%********************************[EQUATION]************************************
\begin{equation}
\mathbf{R}^{m}_{i,j} (\epsilon) = \Theta ( \epsilon_i - || X(i) - X(j) ||
\end{equation}
%%********************************[EQUATION]************************************
where $\quad i,j=1,\dots,N$, $\epsilon$ is a threshold distance, 
$|| \cdotp ||$ a norm, and $\Theta(\cdotp)$ is the Heaviside 
function (i.e. $\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) 
(Fig~\ref{fig:mrp}) \cite{eckmann1987, marwan2007,marwan2015}.
RP is also characterised with a line of identity (LOI) which is a  
diagonal line due to $ R_{i,j}=1 (i,j=1,\dots,N)$. 
%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/methods/rp/pdf/rp}
    \caption{
	{\bf Recurrence Plots.} 
	(A) State space of the Lorenz system with controlling parameters 
	($\rho=28, \sigma=10, \beta=8/3$). 
	A point, $j$, in trajectory $X()$ which falls into the neighborhood 
	(black circle) of a given point at $i$ is a recurrent point and is 
	represented as a black dot in the recurrence plot at location 
	$(i, j)$ or white otherwise.
	(B) Recurrence plot using the 
	three components of the Lorenz system and the RP with no embeddings 
	and threshold $\epsilon=5$.
	This figure is adapted from \cite{marwan2015}.
	Code and data to reproduce the figure is available in \cite{srep2020}.
	}
    \label{fig:mrp}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\subsection*{Structures of Recurrence Plots}
Pattern formations in the RPs can be designated either 
as topology for large-scale patterns or texture for small-scale patterns.
In the case of topology, the following pattern formations are commonly presented:
(i) homogeneous where uniform recurrence points are spread in the RP e.g., 
uniformly distributed noise (Fig~\ref{fig:rp2}A), 
(ii) periodic and quasi-periodic systems where diagonal lines and 
checkerboard structures represent oscillating systems, e.g., sinusoidal 
signals (Fig~\ref{fig:rp2}B), 
(iii) drift where paling or darkening recurrence points away from 
the LOI is caused by drifting systems, 
e.g., logistic map (Fig~\ref{fig:rp2}C), and
(iv) disrupted where recurrence points are presented white areas or 
bands that indicate abrupt changes in the dynamics, e.g. Brownian motion 
(Fig~\ref{fig:rp2}D) \cite{eckmann1987, marwan2015}.
Texture patterns in RPs can be categorised as:
(i) single or isolated recurrence points that represent rare occurring states, 
and do not persist for any time or fluctuate heavily,
(ii) dots forming diagonal lines where the length of the small-scale parallel 
lines in the diagonal are related to the ratio of determinism or predictability 
in the dynamics of the system, and
(iii) dots forming vertical and horizontal lines where the length of the 
lines represent a time length where a state does not change or change very 
slowly and these patterns formation represent discontinuities in the signal, 
and (iv) dots clustering to inscribe rectangular regions which are related 
to laminar states or singularities \cite{marwan2015}.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/methods/rpsp/pdf/rpsp}
    \caption{
	{\bf Patterns in Recurrence Plots.} 
	Time-series with its respective recurrence plot for:
	(A) uniformly distributed noise,
	(B) super-positionet harmonic oscillation 
	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
	(C) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
	with a linearly increase term ($0.01*i$), and
	(D) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
	This figure is adapted from \cite{marwan2015}.
	Code and data to reproduce the figure is available in \cite{srep2020}.
	}
    \label{fig:rp2}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

@@

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\section*{Experiment} \label{sec:experiment}
We conducted an experiment in the context of human-humanoid imitation (HHI) 
activities where participants were asked to imitate simple horizontal and 
vertical arm movements performed by NAO, a humanoid robot \cite{gouaillier2009}.
Such simple movements were repeated ten times for the participant 
who copied NAO's arm movements in a face-to-face imitation activity.
Also, wearable inertial measurement unit (IMU) sensors were attached 
to the right hand of the participant and to the left hand of the robot 
(Figure~\ref{fig:hri} A,C). Data were then collected with four NeMEMSi 
IMU sensors with sampling rate of 50Hz provinding tri-axial data of the 
accelerometer, gyroscope and magnetometer sensors and quaternions 
\cite{Comotti2014}.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
  \centering
\includegraphics[width=1.0\textwidth]{figures/experiment/hri/pdf/hri}
    \caption{
	{\bf Human-humanoid imitation activities.} 
    		Face-to-face human-humanoid imitation (HHI) activities for 
		(A) HHI of horizontal arm movement, 
		(B) Humanoid horizontal arm movement,
		(C) HHI of vertical arm movement, and 
		(D) Humanoid vertical arm movement.
        }
    \label{fig:hri}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

@@

\subsection*{Human-humanoid imitation activities}
For human-humanoid imitation (HHI) activities four neMEMSi sensors were used,
two of which were attached to the right hand of the participant and the 
other two to the left hand of the humanoid robot.
Then, each participant was asked to imitate repetitions of simple horizontal
and vertical arm movements performed by the humanoid robot in the following 
conditions:
(i) ten repetitions of horizontal arm movement at normal (HN) 
	and faster (HF) speed (Figure~\ref{fig:hri} A), and
(ii) ten repetitions of vertical arm movement at normal (VN) 
	and faster (VF) speed (Figure~\ref{fig:hri} C).
The normal and faster speed of arm movements is defined by the duration 
in number of samples of one repetition of NAO's arm movements. 
We select NAO's arm movements duration to distinguish between normal and 
faster arm movements as NAO's movements have less variation 
for such repetitive movements. 
The duration for one repetition of the horizontal 
arm movement at normal speed, HN, is 5 seconds considering 
that each repetition last around 250 samples.
For horizontal arm movement at faster speed, HF, each repetition were performed 
in 2 seconds which correspond to 90 samples of data.
The vertical arm movement at normal speed, VN, were performed  in 6 seconds 
which is around 300 samples of data. For vertical arm movement at 
faster speed, VF, each repetition lasts about 2.4 seconds which correspond 
to 120 samples of data.
To visualise the distinction between normal and faster speed for horizontal 
and vertical arm movements, Fig~\ref{fig:sts} shows smoothed time series 
for axes Z and Y of the gyroscope sensors with four window lengths: 
2-sec (100-samples), 5-sec (250-samples), 10-sec (500-samples) 
and 15-sec (750-samples).
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht] 
\centering
\includegraphics[width=1.0\textwidth]{figures/experiment/sts/pdf/sts}
    \caption{
	{\bf Time series duration of horizontal and vertical arm movements.} 
    		Time series of smoothed data from gyroscope sensor 
		for different speed arm movements performed by NAO: 
		(A) Horizontal Normal arm movement (HN), 
		(B) Horizontal Faster arm movement (HF),
		(C) Vertical Normal arm movement (VN) and 
		(D) Vertical Faster arm movement (VF).
		Additionally, 
		(A) shows window sizes for 2-seconds (100 samples), 
		5-seconds (250 samples), 10-seconds (500 samples) 
		and 15-seconds (750 samples)
		which are also presented in (B), (C) and (D).
		Code and data to reproduce the figure is available in \cite{srep2020}.
        }
	\label{fig:sts}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

\subsection*{Data collection from inertial measurement units} 
\label{sec:experiment:subsec:imu}
To give insight to the research questions, 
we considered various conditions of time series collected for this work: 
\begin{itemize}
\item Three levels of smoothness for normalised data 
	(e.g., sg0zmuv, sg1zmuv and sg2zmuv) where sg 
	stands for Savitzky-Golay filter with two different filter lengths (29 and 159) 
	and the same polynomial degree of 5 using the function \texttt{sgolay(p,n,m)} \cite{Rsignal}
	and zmuv is zero mean unit variance.
\item four arm movement activities with two velocities: 
	horizontal normal (HN), horizontal faster (HF), 
	vertical normal (VN), and vertical faster (VF), and
\item four window lengths: 2-sec (100 samples), 5-sec (250 samples), 
	10-sec (500 samples) and 15-sec (750 samples).
\end{itemize}
%After the data collection, raw time series were windowed, normalised and smoothed. 
Due to space limitations and to have simple visualisation, 
we only present 10-sec (500 samples) window length time series for 
three participants (p01, p01 and p03) performing horizontal 
arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig:ts}).
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/timeseries/pdf/fig1}
    	\caption{ 
	{\bf Time series for horizontal and vertical arm movements.}
		(A/D) raw-normalised (sg0zmuv), 
		(B/E) normalised-smoothed 1 (sg1zmuv) and
		(C/F) normalised-smoothed 2 (sg2zmuv).
		Time series are only for three participants 
		($p01$, $p02$, and $p03$) 
		for horizontal and vertical arm movements in normal 
		and faster velocity (HN, HF, VN, VF) 
		with the normalised GyroZ or GyroY axis 
		and with one sensor attached to the participant (HS01) 
		and other sensor attached to the robot (RS01).	
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:ts}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

@@

%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/caoami/pdf/fig2}
	\caption{
	{\bf Box plots of minimum embedding parameters.} 
	Box plots of (A) minimum embedding dimensions 
	and (B) first minimum AMI values for 
	Horizontal Normal (HN), Horizontal Faster (HF),
	Vertical Normal (VN) and Vertical Faster (VF)
	with sensors attached to participants (HS01) and
	sensor attached to robot (RS01).
	Minimum embedding parameters are for twenty participants 
	($p01$ to $p20$) with three smoothed signals 
	(sg0: sg0zmuvGyroZ, sg1: sg1zmuvGyroZ and sg2: sg2zmuvGyroZ)
	and window length of 10-sec (500 samples).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:cao_ami}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

@@

%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rss/pdf/fig3}
\caption{
	{\bf RSSs for horizontal arm movements.}
	Reconstructed state spaces for time series for p01 of Figure \ref{fig:ts}.
	Reconstructed state spaces were computed with 
	embedding parameters 
	$\overline{m}_0=6$, $\overline{\tau}_0=8$.
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:rss_aHw10}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{figures/rss/pdf/fig4}
    \caption{
	{\bf RSSs for vertical arm movements.}
	Reconstructed state spaces for time series for p01 of Figure \ref{fig:ts}.
	Reconstructed state spaces were computed with 
	embedding parameters
	$\overline{m}_0=6$, $\overline{\tau}_0=8$.
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:rss_aVw10}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

>>>>

%*******************************************************************************
\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
%\graphicspath{{../}} %goes to path: docs/

@@

\includegraphics[width=1.0\textwidth]{rss.pdf}

@@

As an example of the use of $E_1(m)$ and $E_2(m)$ values, we consider two time 
series: the solution for the $x$ variable of the Lorenz system 
(Fig~\ref{fig:e1e2}E), and a Gaussian noise time series with zero mean 
and a variance of one (Fig~\ref{fig:e1e2}F).
We then compute $E_1(m)$ and $E_2(m)$ values for each time series.
The $E_1(m)$ values for the chaotic time series appear to be constant
after the dimension is equal to six.
The determination of six is given that any value of $m$ can be used as they 
are within the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}A).
$E_2(m)$ values, for chaotic time series, are different to one 
(Fig~\ref{fig:e1e2}C), for which, it can be concluded that for the 
chaotic time series the minimum embedding dimension the time series 
comes from a deterministic signal. With regard to the noise time series,  
$E_1(m)$ values appeared to be constant when $m$ is close to thirteen, 
which is defined by the threshold of $1\pm0.05$ (Fig~\ref{fig:e1e2}B).
$E_1(m)$ values then indicate the minimum embedding dimension of the 
noisy time series is thirteen, however all of the $E_2(m)$ values are 
approximately equal to one (Figure~\ref{fig:e1e2}D) for which it can be 
concluded that noise time series is a stochastic signal.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{cao.pdf}
    \caption{
	{\bf Minimum dimension embedding values with Cao's method.} 
	(A, B) $E_1 (m)$ values and (C, D) $E_2(m)$ values 
	with variations of $\tau$ values from one to twenty
	for (E) chaotic and (F) random time series.
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:e1e2}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

@@

For example, we compute the AMI for two time series:
A) the $x$ solution of the Lorenz system, and
B) a noise time series using a normal distribution with mean zero and 
standard deviation equal to one. From Fig~\ref{fig:amis}, it can then be 
concluded that the amount of knowledge for any noise time series is zero 
for which the first minimum embedding parameter is $\tau_0=1$. 
On the contrary, the first minimum of the AMI for the chaotic time series 
is $\tau_0=17$ which is the value that maximize the independence 
between $x(n)$ and $x(n+\tau)$ in the reconstructed state space 
\cite{bradley2015}.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{ami.pdf}
    \caption{
	{\bf Minimum delay embedding values with AMI's method.} 
    	(A, B) AMI values where its first minimum value in the curve
	is the minimum time delay embedding ($\tau_0$), 
	for (C) a chaotic and (D) noise time series.
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:amis}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------
Similarly as Cao's algorithm negatives, AMI's algorithm is not an
exception for negatives, which are worthwhile to mention for further 
investigations.
For instance, (i) is not clear why the choose of the first minimum of the AMI 
is the minimum delay embedding parameter \cite{kantz2003} and 
(ii) the probability distribution of the AMI function is computed
with the use of histograms which depends on a heuristic choice of number of bins
for which AMI depends on partitioning \cite{garcia2005e71}.

@@

For the creation of a recurrence plot based on time series 
$\{ \boldsymbol{x}_n \}$, it is first computed the state space 
reconstruction with uniform time-delay embedding 
$X(i)=\{ \boldsymbol{ \tilde{x} }_n, \dots,  
\boldsymbol{ \tilde{x} }_{n -(m-1)\tau} \}$
where $i=1,\dots,N$, $N$ is the number of considered states of $X(i)$ 
and $X(i) \in \mathbb{R}^m$ \cite{eckmann1987}.
The recurrence plot is therefore a two-dimensional $N \times N$ square 
matrix, $\mathbf{R}$, where a black dot is placed at $(i,j)$ 
whenever $X(i)$ is sufficiently close to $X(j)$: 
%%********************************[EQUATION]************************************
\begin{equation}
\mathbf{R}^{m}_{i,j} (\epsilon) = \Theta ( \epsilon_i - || X(i) - X(j) ||
\end{equation}
%%********************************[EQUATION]************************************
where $\quad i,j=1,\dots,N$, $\epsilon$ is a threshold distance, 
$|| \cdotp ||$ a norm, and $\Theta(\cdotp)$ is the Heaviside 
function (i.e. $\Theta(x)=0$, if $x<0$, and $\Theta(x)=1$ otherwise) 
(Fig~\ref{fig:mrp}) \cite{eckmann1987, marwan2007,marwan2015}.
RP is also characterised with a line of identity (LOI) which is a  
diagonal line due to $ R_{i,j}=1 (i,j=1,\dots,N)$. 
%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{rp}
    \caption{
	{\bf Recurrence Plots.} 
	(A) State space of the Lorenz system with controlling parameters 
	($\rho=28, \sigma=10, \beta=8/3$). 
	A point, $j$, in trajectory $X()$ which falls into the neighborhood 
	(black circle) of a given point at $i$ is a recurrent point and is 
	represented as a black dot in the recurrence plot at location 
	$(i, j)$ or white otherwise.
	(B) Recurrence plot using the 
	three components of the Lorenz system and the RP with no embeddings 
	and threshold $\epsilon=5$.
	This figure is adapted from \cite{marwan2015}.
	Code and data to reproduce the figure is available in \cite{srep2020}.
	}
    \label{fig:mrp}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\subsection*{Structures of Recurrence Plots}
Pattern formations in the RPs can be designated either 
as topology for large-scale patterns or texture for small-scale patterns.
In the case of topology, the following pattern formations are commonly presented:
(i) homogeneous where uniform recurrence points are spread in the RP e.g., 
uniformly distributed noise (Fig~\ref{fig:rp2}A), 
(ii) periodic and quasi-periodic systems where diagonal lines and 
checkerboard structures represent oscillating systems, e.g., sinusoidal 
signals (Fig~\ref{fig:rp2}B), 
(iii) drift where paling or darkening recurrence points away from 
the LOI is caused by drifting systems, 
e.g., logistic map (Fig~\ref{fig:rp2}C), and
(iv) disrupted where recurrence points are presented white areas or 
bands that indicate abrupt changes in the dynamics, e.g. Brownian motion 
(Fig~\ref{fig:rp2}D) \cite{eckmann1987, marwan2015}.
Texture patterns in RPs can be categorised as:
(i) single or isolated recurrence points that represent rare occurring states, 
and do not persist for any time or fluctuate heavily,
(ii) dots forming diagonal lines where the length of the small-scale parallel 
lines in the diagonal are related to the ratio of determinism or predictability 
in the dynamics of the system, and
(iii) dots forming vertical and horizontal lines where the length of the 
lines represent a time length where a state does not change or change very 
slowly and these patterns formation represent discontinuities in the signal, 
and (iv) dots clustering to inscribe rectangular regions which are related 
to laminar states or singularities \cite{marwan2015}.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{rpsp}
    \caption{
	{\bf Patterns in Recurrence Plots.} 
	Time-series with its respective recurrence plot for:
	(A) uniformly distributed noise,
	(B) super-positionet harmonic oscillation 
	($sin( \frac{1}{5}*t) * sin( \frac{5}{100}*t) $),
	(C) drift logistic map ($x_{i+1} = 4 x_i (1- x_i) $) corrupted 
	with a linearly increase term ($0.01*i$), and
	(D) disrupted brownian motion  ($x_{i+1} = x_i + 2*rnorm(1) $).
	This figure is adapted from \cite{marwan2015}.
	Code and data to reproduce the figure is available in \cite{srep2020}.
	}
    \label{fig:rp2}
\end{figure}
%%---------------------------------(FIGURE)-------------------------------------

@@

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\section*{Experiment} \label{sec:experiment}
We conducted an experiment in the context of human-humanoid imitation (HHI) 
activities where participants were asked to imitate simple horizontal and 
vertical arm movements performed by NAO, a humanoid robot \cite{gouaillier2009}.
Such simple movements were repeated ten times for the participant 
who copied NAO's arm movements in a face-to-face imitation activity.
Also, wearable inertial measurement unit (IMU) sensors were attached 
to the right hand of the participant and to the left hand of the robot 
(Figure~\ref{fig:hri} A,C). Data were then collected with four NeMEMSi 
IMU sensors with sampling rate of 50Hz provinding tri-axial data of the 
accelerometer, gyroscope and magnetometer sensors and quaternions 
\cite{Comotti2014}.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
  \centering
\includegraphics[width=1.0\textwidth]{hri}
    \caption{
	{\bf Human-humanoid imitation activities.} 
    		Face-to-face human-humanoid imitation (HHI) activities for 
		(A) HHI of horizontal arm movement, 
		(B) Humanoid horizontal arm movement,
		(C) HHI of vertical arm movement, and 
		(D) Humanoid vertical arm movement.
        }
    \label{fig:hri}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

@@

\subsection*{Human-humanoid imitation activities}
For human-humanoid imitation (HHI) activities four neMEMSi sensors were used,
two of which were attached to the right hand of the participant and the 
other two to the left hand of the humanoid robot.
Then, each participant was asked to imitate repetitions of simple horizontal
and vertical arm movements performed by the humanoid robot in the following 
conditions:
(i) ten repetitions of horizontal arm movement at normal (HN) 
	and faster (HF) speed (Figure~\ref{fig:hri} A), and
(ii) ten repetitions of vertical arm movement at normal (VN) 
	and faster (VF) speed (Figure~\ref{fig:hri} C).
The normal and faster speed of arm movements is defined by the duration 
in number of samples of one repetition of NAO's arm movements. 
We select NAO's arm movements duration to distinguish between normal and 
faster arm movements as NAO's movements have less variation 
for such repetitive movements. 
The duration for one repetition of the horizontal 
arm movement at normal speed, HN, is 5 seconds considering 
that each repetition last around 250 samples.
For horizontal arm movement at faster speed, HF, each repetition were performed 
in 2 seconds which correspond to 90 samples of data.
The vertical arm movement at normal speed, VN, were performed  in 6 seconds 
which is around 300 samples of data. For vertical arm movement at 
faster speed, VF, each repetition lasts about 2.4 seconds which correspond 
to 120 samples of data.
To visualise the distinction between normal and faster speed for horizontal 
and vertical arm movements, Fig~\ref{fig:sts} shows smoothed time series 
for axes Z and Y of the gyroscope sensors with four window lengths: 
2-sec (100-samples), 5-sec (250-samples), 10-sec (500-samples) 
and 15-sec (750-samples).
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht] 
\centering
\includegraphics[width=1.0\textwidth]{sts}
    \caption{
	{\bf Time series duration of horizontal and vertical arm movements.} 
    		Time series of smoothed data from gyroscope sensor 
		for different speed arm movements performed by NAO: 
		(A) Horizontal Normal arm movement (HN), 
		(B) Horizontal Faster arm movement (HF),
		(C) Vertical Normal arm movement (VN) and 
		(D) Vertical Faster arm movement (VF).
		Additionally, 
		(A) shows window sizes for 2-seconds (100 samples), 
		5-seconds (250 samples), 10-seconds (500 samples) 
		and 15-seconds (750 samples)
		which are also presented in (B), (C) and (D).
		Code and data to reproduce the figure is available in \cite{srep2020}.
        }
	\label{fig:sts}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

\subsection*{Data collection from inertial measurement units} 
\label{sec:experiment:subsec:imu}
To give insight to the research questions, 
we considered various conditions of time series collected for this work: 
\begin{itemize}
\item Three levels of smoothness for normalised data 
	(e.g., sg0zmuv, sg1zmuv and sg2zmuv) where sg 
	stands for Savitzky-Golay filter with two different filter lengths (29 and 159) 
	and the same polynomial degree of 5 using the function \texttt{sgolay(p,n,m)} \cite{Rsignal}
	and zmuv is zero mean unit variance.
\item four arm movement activities with two velocities: 
	horizontal normal (HN), horizontal faster (HF), 
	vertical normal (VN), and vertical faster (VF), and
\item four window lengths: 2-sec (100 samples), 5-sec (250 samples), 
	10-sec (500 samples) and 15-sec (750 samples).
\end{itemize}
%After the data collection, raw time series were windowed, normalised and smoothed. 
Due to space limitations and to have simple visualisation, 
we only present 10-sec (500 samples) window length time series for 
three participants (p01, p01 and p03) performing horizontal 
arm movements (axis GyroZ) and vertical arm movements (axis GyroY) (Figs \ref{fig:ts}).
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{fig1}
    	\caption{ 
	{\bf Time series for horizontal and vertical arm movements.}
		(A/D) raw-normalised (sg0zmuv), 
		(B/E) normalised-smoothed 1 (sg1zmuv) and
		(C/F) normalised-smoothed 2 (sg2zmuv).
		Time series are only for three participants 
		($p01$, $p02$, and $p03$) 
		for horizontal and vertical arm movements in normal 
		and faster velocity (HN, HF, VN, VF) 
		with the normalised GyroZ or GyroY axis 
		and with one sensor attached to the participant (HS01) 
		and other sensor attached to the robot (RS01).	
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:ts}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

@@

%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{fig2}
	\caption{
	{\bf Box plots of minimum embedding parameters.} 
	Box plots of (A) minimum embedding dimensions 
	and (B) first minimum AMI values for 
	Horizontal Normal (HN), Horizontal Faster (HF),
	Vertical Normal (VN) and Vertical Faster (VF)
	with sensors attached to participants (HS01) and
	sensor attached to robot (RS01).
	Minimum embedding parameters are for twenty participants 
	($p01$ to $p20$) with three smoothed signals 
	(sg0: sg0zmuvGyroZ, sg1: sg1zmuvGyroZ and sg2: sg2zmuvGyroZ)
	and window length of 10-sec (500 samples).
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:cao_ami}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

@@

%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{fig3}
\caption{
	{\bf RSSs for horizontal arm movements.}
	Reconstructed state spaces for time series for p01 of Figure \ref{fig:ts}.
	Reconstructed state spaces were computed with 
	embedding parameters 
	$\overline{m}_0=6$, $\overline{\tau}_0=8$.
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:rss_aHw10}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{fig4}
    \caption{
	{\bf RSSs for vertical arm movements.}
	Reconstructed state spaces for time series for p01 of Figure \ref{fig:ts}.
	Reconstructed state spaces were computed with 
	embedding parameters
	$\overline{m}_0=6$, $\overline{\tau}_0=8$.
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:rss_aVw10}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

\subsection*{Recurrences Plots}
Using the average embedding parameters 
($\overline{m}_0=6$, $\overline{\tau}_0=8$) 
and an recurrence threshold of $\epsilon=1$.
As our interest is for dynamical transitions, 
there is little importance on the selection of $\epsilon$ which in this case
is 1. 
Recurrence Plots (RP) were computed for horizontal (figures~\ref{fig:rp_aH}) and 
vertical (figures~\ref{fig:rp_aV}) arm movements.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{rp_aH}
\caption{
	{\bf RPs for horizontal arm movements.}	
	Recurrence plots were computed with 
	embedding parameters 
	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:rp_aH}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
\centering
\includegraphics[width=1.0\textwidth]{rp_aV}
\caption{
	{\bf RPs for vertical arm movements.}	
	Recurrence plots were computed with 
	$\overline{m}_0=6$, $\overline{\tau}_0=8$, and $\epsilon=1$.
	Code and data to reproduce the figure is available in \cite{srep2020}.
        }
    \label{fig:rp_aV}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------
